---
output: 
  html_document:
    fig_caption: yes
    css: www/webstyles.css
    anchor_sections: FALSE
    includes: 
      in_header: "header_manual.html" 
      after_body: "footer.html"
    self_contained: yes
title: "NETN water summary"
params:
  park: MABI # 4-letter park code
  year_current: 2024 # current year of data
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = "h", warning = FALSE, message = FALSE, fig.align = 'left')
```

```{r include = F, results = 'hide'}
library(tidyverse)
library(grid)
library(gridExtra)
library(kableExtra)
#library(plotly)
```

```{r }
# park = "MABI"; yr_start = 2006; yr_current = 2023;
parkcode <- child_env$park
parkname <- child_env$park_df$UnitName[child_env$park_df$UnitCode == parkcode]
```

##  `r park` {.tabset}

```{r}
yr5 = yr_current - 4
drgt_cumppt_cap1 <- paste0("Weekly drought condition for ", yr_current, " based on proportion of a park's predominant county in each drought level using the <a href='https://droughtmonitor.unl.edu/'>U.S. Drought Monitor</a> (left). Cumulative annual precipitation in ", yr_current, " by month compared to the 20th century baseline (right). Total monthly precipitation data are derived from  <a href='https://www.ncei.noaa.gov/access/metadata/landing-page/bin/iso?id=gov.noaa.ncdc:C00332'>NOAA monthly NClimGrid gridded climate data</a> extracted for a given park's geographic centroid.") 

anom_comp_cap1 <- paste0("Current (", yr_current, ") monthly climate data (solid) compared to historic climate baselines (left). Climate baselines include minimum and maximum (light grey), 95% distribution (grey), and middle 50% distribution (dark grey) of values recorded from 1901 to 2000, along with the baseline average (black line). Monthly deviations from historic baselines (right) in ", yr_current, " with positive (red) bars indicating above average deviations, and negative (blue) bars indicating below average deviations. Historic baselines are based on <a href='https://www.ncei.noaa.gov/products/land-based-station/us-climate-normals'>NOAA 20th century normals</a> for a given park's geographic centroid. Current monthly data are derived from <a href='https://www.ncei.noaa.gov/access/metadata/landing-page/bin/iso?id=gov.noaa.ncdc:C00332'>NOAA NClimGrid </a> gridded climate dataset extracted for a given park's geographic centroid.") 

anom_cap5 <- paste0("Monthly deviations from historic baselines for past 5 years (", yr5, ":", yr_current, ") with positive (red) bars, indicated above average deviations, and negative (blue) bars indicating below average deviations. Historic baselines are based on <a href='https://www.ncei.noaa.gov/products/land-based-station/us-climate-normals'>NOAA 20th century normals</a> for a given park's geographic centroid. 5-year  (", yr5, ":", yr_current, ") monthly data are derived from <a href='https://www.ncei.noaa.gov/access/metadata/landing-page/bin/iso?id=gov.noaa.ncdc:C00332'>NOAA NClimGrid </a> gridded climate dataset extracted for a given park's geographic centroid.")

cum_cap5 <- paste0("Cumulative annual precipitation by month for past 5 years (", yr5, ":", yr_current, " ) compared to the 20th century baseline. Total monthly precipitation data are derived from  <a href='https://www.ncei.noaa.gov/access/metadata/landing-page/bin/iso?id=gov.noaa.ncdc:C00332'>NOAA monthly NClimGrid gridded climate data</a> extracted for a given park's geographic centroid.")

drgt_cap5 <- paste0("Last 5 years (", yr5, ":", yr_current, ") of monthly average temperature (top), monthly total precipitation (middle), and weekly drought condition (bottom) based on proportion of a park's predominant county by drought level using the <a href='https://droughtmonitor.unl.edu/'>U.S. Drought Monitor</a>.") 

temp_capT <- paste0("5-year rolling average annual minimum, maximum, and average temperature trends from 1900 to ", yr_current, " by month derived from <a href='https://www.ncei.noaa.gov/access/metadata/landing-page/bin/iso?id=gov.noaa.ncdc:C00332'>NOAA NClimGrid </a> gridded climate dataset extracted for a given park's geographic centroid.")

ppt_capT <- paste0("5-year rolling average annual total precipitation trends from 1900 to ", yr_current, " derived from <a href='https://www.ncei.noaa.gov/access/metadata/landing-page/bin/iso?id=gov.noaa.ncdc:C00332'>NOAA NClimGrid </a> gridded climate dataset extracted for a given park's geographic centroid.")

anom_capT <- paste0("Monthly deviations from historic baselines, with positive (red) bars, indicated above average deviations, and negative (blue) bars indicating below average deviations for 1900 to ", yr_current, ". Historic baselines are based on <a href='https://www.ncei.noaa.gov/products/land-based-station/us-climate-normals'>NOAA 20th century normals</a> for a given park's geographic centroid. Monthly data are derived from <a href='https://www.ncei.noaa.gov/access/metadata/landing-page/bin/iso?id=gov.noaa.ncdc:C00332'>NOAA NClimGrid </a> gridded climate dataset extracted for a given park's geographic centroid.")

drgt_capT <- paste0("Long-term trends (2000:", yr_current, ") in weekly drought condition based on proportion of a park's predominant county by drought level using the <a href='https://droughtmonitor.unl.edu/'>U.S. Drought Monitor</a>. Note that drought data are only available from 2000 to present.") 

```

### Current Year {.tabset}
<h1>Current vs. 20th Century Baseline</h1>
<details open><summary class='drop'>Temperature</summary>
```{r fig.height=7, fig.width=13, results = 'asis', fig.cap = anom_comp_cap1}
temp_comp <- plotClimComps(park = parkcode, years = yr_current, palette = 'red', parameter = c('tmean'),
                           legend_position = 'bottom', plot_title = F, units = "eng", 
                           layers = c("points", "lines"), legend_row = 2)
temp_anom <- plotClimAnom(park = parkcode, years = yr_current, parameter = 'tmean',
                          legend_position = 'bottom', units = "eng")
grid.arrange(temp_comp, temp_anom, nrow = 1)

#cat(anom_comp_cap1, "\n\n")
```

</details>
<br>
<details open><summary class='drop'>Precipitation</summary>
```{r fig.height=7, fig.width=13, results = 'asis', fig.cap = anom_comp_cap1}
ppt_comp <- plotClimComps(park = park,years= yr_current, palette = 'blue', parameter = c('ppt'),
                           legend_position = "bottom", plot_title = F, units = "eng", 
                          layers = c("points", "lines"), legend_row = 2)
ppt_anom <- plotClimAnom(park = parkcode, years = yr_current, parameter = 'ppt_pct',
                          legend_position = "bottom", palette = c("#1e90ff", "#7F6748"))
grid.arrange(ppt_comp, ppt_anom, nrow = 1)

#cat(anom_comp_cap1, "\n\n")
```
</details>
<br>
<details open><summary class='drop'>Drought Condition</summary>
```{r fig.height=7, fig.width=13, results = 'asis', fig.cap = drgt_cumppt_cap1}
drgt <- plotClimDrought(park = parkcode, years = yr_current, 
                        legend_position = 'bottom', legend_row = 2, 
                        end_date = child_env$end_date)
pptcum <- plotClimCumPrecip(park = parkcode, years = yr_current, 
                            legend_position = "bottom", palette = "#1e90ff", 
                            layers = c("points", "lines"), legend_row = 2) 
cat("\n\n")
grid.arrange(drgt, pptcum, nrow = 1)
cat("\n\n")

#cat("\n", drgt_cumppt_cap1, "\n\n")
```
</details>
<br>

### 5-year Trends {.tabset}
<h1>5-year vs. 20th Century Baseline</h1>
<details open><summary class='drop'>Temperature</summary>
```{r fig.height=6, fig.width=13, results = 'asis', fig.cap = anom_cap5}
yr1 = yr_current - 4
plotClimAnom(park = parkcode, years = yr1:yr_current, parameter = 'tmean', 
             legend_position = "bottom", gridlines = "grid_x", units = "eng")
#cat(anom_cap5, "\n\n")
```
</details>
<br>

<details open><summary class='drop'>Total Precipitation</summary>
```{r fig.height=6, fig.width=13, results = 'asis', fig.cap = anom_cap5}
plotClimAnom(park = parkcode, years = yr1:yr_current, parameter = 'ppt_pct', 
             legend_position = "bottom", gridlines = "grid_x", palette = c("#1e90ff", "#7F6748"))
#cat(anom_cap5, "\n\n")
```
</details>
<br>

<details open><summary class='drop'>Drought Condition</summary>
```{r fig.height=7.5, fig.width=12.1, results = 'asis', fig.cap = drgt_cap5}
temp <- plotClimTrend(park = parkcode, years = yr1:yr_current, units = "eng",
                      gridlines = "grid_x", parameter = "tmax", 
                      facet_param = F, layers = "lines", 
                      palette = c("red"), line_width = 1, x_pad = c(0,29)) + 
  theme(axis.text.x = element_blank(), 
        axis.title.y = element_text(size = 9), 
        plot.margin = margin(l = 5, r = 8, b = 7, t = 2))

ppt <- plotClimTrend(park = parkcode, years = yr1:yr_current, units = "eng",
                     gridlines = 'grid_x', parameter = 'ppt',
                     facet_param = F, layers = "bar",
                     palette = "#1e90ff", line_width = 1, x_pad = c(0,29)) +
  theme(axis.text.x = element_blank(), 
        axis.title.y = element_text(size = 9), 
        plot.margin = margin(l = 5, r = 8, b = 2))

drgt <- plotClimDrought(park = parkcode, years = yr1:yr_current, gridlines = "grid_x", 
                legend_position = "bottom", end_date = child_env$end_date, x_pad = c(0.01,0)) +
        theme(axis.title.y = element_text(size = 9))

grid.arrange(ggplotGrob(temp), ggplotGrob(ppt), ggplotGrob(drgt), 
             nrow = 3, ncol = 1, heights = c(0.4, 0.4, 1.5))

#cat(drgt_cap5, "\n\n")
```
</details>
</br>


<details open><summary class='drop'>Cumulative Precipitation</summary>
```{r fig.height=7, fig.width=13, results = 'asis', fig.cap = cum_cap5}
plotClimCumPrecip(park = parkcode, years = yr1:yr_current, legend_position = "bottom", 
                  units = "eng", palette = "#1e90ff", layers = c("points", "lines"))

#cat(cum_cap5, "\n\n")

```
</details>
</br>

### Long-term Trends {.tabset}
<h1>Historic Trends 1900:2024</h1>
<details open><summary class='drop'>Temperature Trends</summary>
```{r fig.height=6.5, fig.width=12.05, results = "asis", fig.cap = temp_capT}
plotClimTrend(park = parkcode, years = 1895:yr_current, 
              parameter = c("tmin", "tmax", "tmean"), facet_param = F,
              legend_position = "bottom", plot_se = T,
              layers = "rollavg", gridlines = "grid_x", units = "eng", 
              palette = c("dimgrey", "#EC6128", "#3B70B0"))
#cat(temp_capT, "\n\n")

```
</details>
<br>

<details open><summary class='drop'>Temperature Anomalies</summary>
```{r fig.height=6.5, fig.width=13, results = 'asis', fig.cap = anom_capT}
plotClimAnom(park = parkcode, years = 1900:yr_current, 
              parameter = c("tmean"), legend_position = "bottom", gridlines = "grid_x", units = "eng")

#cat(anom_capT, "\n\n")
```
</details>
<br>

<details open><summary class='drop'>Precipitation Trends</summary>
```{r fig.height=6.5, fig.width=12.05, results = 'asis', fig.cap = ppt_capT}
plotClimTrend(park = parkcode, years = 1895:yr_current,
              parameter = c("ppt"), facet_param = F, plot_se = T,
              palette = "#3B70B0", legend_position = "bottom", 
              layers = c("rollavg"), gridlines = "grid_x", units = "eng")

#cat(ppt_capT, "\n\n")
```
</details>
<br>

<details open><summary class='drop'>Precipitation Anomalies</summary>
```{r fig.height=6.5, fig.width=13, results = 'asis', fig.cap = anom_capT}
plotClimAnom(park = parkcode, years = 1900:yr_current, gridlines = "grid_x", 
              parameter = c("ppt_pct"), legend_position = "bottom", palette = c("#1e90ff", "#7F6748"))

#cat(anom_capT, "\n\n")
```
</details>
<br>


<details open><summary class='drop'>Drought Conditions</summary>
```{r fig.height=6.5, fig.width=12.1, results = 'asis', fig.cap = drgt_capT}
# temp <- plotClimTrend(park = parkcode, years = 2000:yr_current, units = "eng",
#                       gridlines = "grid_x", parameter = "tmax", 
#                       facet_param = F, layers = "lines", 
#                       palette = c("red"), line_width = 1, x_pad = c(0,29)) + 
#   theme(axis.text.x = element_blank(), 
#         axis.title.y = element_text(size = 9), 
#         plot.margin = margin(l = 15, r = 12, b = 7, t = 2))
# 
# ppt <- plotClimTrend(park = parkcode, years = 2000:yr_current, units = "eng",
#                      gridlines = 'grid_x', parameter = 'ppt',
#                      facet_param = F, layers = "bar",
#                      palette = "#1e90ff", line_width = 1, x_pad = c(0,29)) +
#   theme(axis.text.x = element_blank(), 
#         axis.title.y = element_text(size = 9), 
#         plot.margin = margin(l = 7, r = 12, b = 2))

#drgt <- 
  plotClimDrought(park = parkcode, years = 2000:yr_current, gridlines = "grid_x", 
                legend_position = "bottom", end_date = child_env$end_date, x_pad = c(0.01,0)) +
        theme(axis.title.y = element_text(size = 9))

# grid.arrange(ggplotGrob(temp), ggplotGrob(ppt), ggplotGrob(drgt), 
#              nrow = 3, ncol = 1, heights = c(0.4, 0.4, 1.5))

```
</details>
<br>

### Table of Records {.tabset}
```{r results = "asis"}
cat(paste0("Table of top 5 monthly records for ", parkname, ". Cells include the year for each record by month, with years >= 2020 denoted in red. Hovering over the cell will display the value for each record.\n\n"))
cat("\n\n")
cat("\n\n")
sumStatsTable(park = parkcode, years = 1895:yr_current, units = "eng", top_n = 5, kable = T)
cat("\n\n")
```

